Black swans in ecological time series\\ Black swans in population dynamics\\ Heavy-tailed ecological time series\\ Black swans in ecological dynamics\\ Black swan ecological dynamics\\ \ldots

### 2014-10-23 text/ms.tex
Edit an intro paragraph

Black swans are unexpected extreme events with dramatic consequences
\citep{taleb2007,sornette2009}. One of the most striking black swans in ecology
was the asteroid marking the mass extinction of the K-T boundary (REF). Today,
it is the extremeness of climate, in concert with shifts in mean temperature,
that is expected to cause the greatest societal damage \citep{ipcc2012}.
But, while extremes in the physical environment such as wave height, rain
fall, and temperature are present and widely accepted
\citep{gaines1993,katz2005}, it remains unclear the extent to which ecological
systems buffer or suffer from black swans \citep{nunez2012}.

%Existing evidence for black swans or heavy tails in population dynamics is
%mostly anecdotal or for particular systems.
There is compelling anecdotal evidence for ecological black swans, but
systematic evidence has been elusive. \citet{doak2008} reviewed a collection of
field experiment population dynamic surprises and concluded that surprises are
more common than we typically assume. \citet{nunez2012} compiled a list of
major events on geologic and contemporary scales that could be considered
ecological black swans. Other work has identified heavy-tail dynamics for
particular systems. For example, \citet{segura2013} identified heavy tails in
time series of marine microbe populations and \citet{drake2014} identified
heavy tails in time to extinction for experimental waterflea populations.
However, systematic evidence is limited. \citet{keitt1998} tested for power-law
tail distributions in breeding bird abundance, but later work showed the
apparent heavy tails were a result of mixing populations of different species
\citep{allen2001}. A subsequent large scale analysis of the Global Population
Dynamics Database (GPDD) by \citet{halley2002} found little evidence for
heavy-tail distributions in longer abundance time series.

>>>>

Black swans are unexpected extreme events with potentially dramatic consequences
\citep{taleb2007,sornette2009}. One of the most striking black swans in ecology
was the asteroid marking the mass extinction of the K-T boundary (REF). Today,
it is the extremeness of climate, in concert with shifts in mean temperature,
that is expected to cause the greatest societal damage \citep{ipcc2012}.
But, while extremes in the physical environment such as wave height, rain
fall, and temperature are present and widely accepted
\citep{gaines1993,katz2005}, it remains unclear the extent to which ecological
systems buffer or suffer from black swans \citep{nunez2012}.

%Existing evidence for black swans or heavy tails in population dynamics is
%mostly anecdotal or for particular systems.
There is compelling anecdotal evidence for ecological black swans, but
systematic evidence has been elusive. A survey of ecologists indicated that
surprising results in field experiments may be far more common than we assume
\citep{doak2008}. And, a number of major environmental events on geologic and
contemporary time scales could be considered to have caused ecological black
swans \citep{nunez2012}. Other work has detected heavy-tail dynamics for
particular systems. For example, \citet{segura2013} identified heavy tails in
time series of marine microbe populations and \citet{drake2014} identified
heavy tails in time to extinction for experimental waterflea populations.
However, systematic evidence has been limited. What appeared to be heavy tails
in breeding bird abundance \citep{keitt1998}, could also be explained by the
mixing of species in the original dataset \citep{allen2001}. A subsequent
large scale analysis of the Global Population Dynamics Database (GPDD) found
little evidence for heavy-tail distributions in longer abundance time series
\citet{halley2002}.
